---
title: "Class05: Data Vis with ggplot"
author: "Neha (PID:A17567541)"
format: pdf
editor: visual
---
# Graphics systems in R

There are many graphic systems in R for making plots and figures

We have already played a little with **"baseR"**
graphics and the `plot()` function

Today we will start learning about a popular graphics package called `ggplot2()`

This is an add on package. We will need to install it. I install it like I install any package with the `install.packages()` function. 

```{r}
plot(cars)
```
Before I can use the functions from a package, I have to load up the package from my "library". We use `library(ggplot2)` command to load it up. 

```{r}
library(ggplot2)
ggplot(cars)
```

Every ggplot is made up of at least 3 things:
-data (the numbers etc. that will go into your plot)
-aes (how the columns of data map to the plot aesthetics)
-geoms (how the plot actually looks, points, bars, lines, etc.)

```{r}
ggplot(cars) +
  aes (x=speed, y=dist)+ geom_point()
```

For simple plots ggplot is more verbose- it takes more code than base R plot. 

Add some more layers to our ggplot. The + sign needs to be at the end of each line

```{r}
ggplot(cars) +
  aes (x=speed, y=dist)+ 
  geom_point() +
  geom_smooth(method="lm")+
  labs(title="Stopping distance of old cars", subtitle= "A silly example plot")+
  theme_bw()

```
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
nrow(genes)
colnames(genes)
ncol(genes)
table(genes$State)
round(table(genes$State)/nrow(genes)*100,2)
```
```{r}
p<-ggplot(genes)+
  aes(x=Condition1, y=Condition2, col=State)+
  geom_point()
p+ scale_colour_manual(values=c("blue","gray","red"))+
  labs(x="Control(no drug)", y="Drug Treatment",title="Gene Expression Changes Upon Drug Treatment")
```
```{r}
# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```
```{r}
 
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)

```
```{r}
ggplot(gapminder_2007)+
  aes(x=gdpPercap,y=lifeExp)+
  geom_point()
```
```{r}
p<-ggplot(gapminder_2007)+
  aes(x=gdpPercap,y=lifeExp)+
  geom_point(alpha=0.4)
```


```{r}
p +
  aes(color=continent, size=pop)
```
```{r}
p+
  aes(color=pop) + 
geom_point(alpha=0.8)
```

```{r}
p+aes(size=pop)
```



